---
title: "Lab 1"
author: "Sebastian Cadavid-Sanchez"
output: html_notebook
---
# Bayesian ERGMs with Bergm R - Lab 1

Code provided by Prof. Alberto Caimo in Networks 2021 and available at [https://acaimo.github.io/teaching/Bergm/home/](https://acaimo.github.io/teaching/Bergm/home/).

## Dataset

Southern Women (Davis) dataset of 18 women observed over a nine-month period. During that period, various subsets of these women had met in a series of 14 informal social events. The data recorded which women met for which events.

```{r}
library(Bergm)
library(latentnet)
```

```{r}
data(davis)
y <- davis

as.matrix.network.adjacency(y)
```
```{r}
# Plot the bipartite network graph
set.seed(7)
plot(y, 
     displaylabels = TRUE, 
     edge.col = "gray",
     vertex.col = c(rep("pink", 18), rep("skyblue", 14)),
     vertex.sides = c(rep(40, 18), rep(4, 14)), 
     vertex.cex = c(rep(3.7, 18), rep(2, 14)), 
     label.cex = 0.4, 
     label.pos = 5)
```
## Model

```{r}
# Model specification
formula <- y ~ edges + gwb2degree(0.7, fixed = TRUE)

# Prior specification 
prior.mean = c(-2, 0) 
prior.sigma = matrix(c(1, -0.5,
                       -0.5, 2), nrow = 2) # covariance matrix
```

```{r, echo=FALSE}
# Prior visualisation
require(MASS)
require(mvtnorm)

r <- hcl.colors(9, "Blues 3")

set.seed(1)
z <- rmvnorm(50000, mean = prior.mean, sigma = prior.sigma)
k <- kde2d(z[, 1],
           z[, 2],
           n = 50)

image(k, col = r,
      xlab = bquote(theta[.(1)]),
      ylab = bquote(theta[.(2)]))
```

## Parameter estimation

```{r}
set.seed(1)
m1.posterior <- bergm(formula,
                      prior.mean = prior.mean,
                      prior.sigma = prior.sigma,
                      aux.iters = 5000, # network simulation iterations
                      main.iters = 2500, 
                      gamma = 1.4) # ADS parameter
```
```{r}
summary(m1.posterior) # Summary of posterior summary

plot(m1.posterior) # Plot the MCMC output
```

```{r}
# Let's visualise the posterior density
require(MASS)
k <- kde2d(c(m1.posterior$Theta[, 1]), 
           c(m1.posterior$Theta[, 2]), 
           n = 100)

image(k, col = r,
      xlab = bquote(theta[.(1)]),
      ylab = bquote(theta[.(2)]))
```

## Model assessment

```{r}
# Draw a sample of 100 paramater values
# from the estimated posterior distribution
set.seed(1)
m1.sampled.pars <- m1.posterior$Theta[sample.int(100), ]

# Predictive network statistics
pred.formula <- y ~ sociality(nodes = 1:18)

s.y.obs <- summary(pred.formula) # observed
pred.formula.names <- ergm_model(pred.formula, y)$coef.names
n.stats <- length(pred.formula.names)

y.sims <- list() # list for the simulated networks
s.y.sims <- matrix(NA, nrow = 100, ncol = n.stats) # simulated statistics
colnames(s.y.sims) <- pred.formula.names

for (i in 1:100) {
  set.seed(1)
  y.sims[[i]] <- simulate(formula,
                          coef = m1.sampled.pars[i, ], 
                          nsim = 1, 
                          control = control.simulate(MCMC.burnin = 5000))
  
  y <- y.sims[[i]]
  s.y.sims[i, ] <- summary(pred.formula)
}
```

```{r}
boxplot(s.y.sims[, 1:n.stats]) # simulated distributions

lines(s.y.obs, lwd = 3, col = "red") # observed values
```

```{r}
y[, ]["DOROTHY", "E9"] <- 1 # observed
s.y.1 <- summary(formula)
y[, ]["DOROTHY", "E9"] <- 0
s.y.0 <- summary(formula)
delta <- s.y.1 - s.y.0
odds <- exp(m1.sampled.pars %*% delta)
boxplot(odds)
```

