---
title: "Lab 2"
author: "Sebastian Cadavid-Sanchez"
output: html_notebook
---

# Bayesian ERGMs with Bergm R - Lab 2

Code provided by Prof. Alberto Caimo in Networks 2021 and available at [https://acaimo.github.io/teaching/Bergm/home/](https://acaimo.github.io/teaching/Bergm/home/).

## Dataset

The Goodreau’s Faux Mesa High School dataset represents a simulation of an in-school friendship network. The school community on which it is based is in the rural western US, with a student body that is largely Hispanic and Native American.

```{r}
require(Bergm)
```

```{r}
data("faux.mesa.high")
y <- faux.mesa.high
y
```

```{r}
r <- hcl.colors(2, "RdYlBu")
set.seed(7)
par(mfrow = c(1, 2)) 
plot(y, 
     vertex.cex = 1,
     edge.col = "grey",
     vertex.col = r[factor(y %v% "Sex")])

legend("topright", 
       pt.bg  = unique(r[factor(y %v% "Sex")]), 
       pt.cex = 1.2,
       pch    = 21, 
       legend = levels(factor(y %v% "Sex")), 
       title  = "Sex")

r <- hcl.colors(5, "Zissou 1")
set.seed(7)
plot(y, 
     vertex.cex = 1,
     edge.col = "grey",
     vertex.col = r[factor(y %v% "Race")])

legend("topright", 
       pt.bg  = unique(r[factor(y %v% "Race")]), 
       pt.cex = 1.2,
       pch    = 21, 
       legend = levels(factor(y %v% "Race")), 
       title  = "Race")


```

## Model

```{r}
# Model specification
formula <- y ~ edges +
               nodematch("Sex") +
               nodematch("Race") +
               gwdegree(0.7, fixed = TRUE)

# Prior specification
prior.mean = c(-4, 2, 2, 0)
prior.sigma = diag(2, 4)
```

## Parameter estimation

```{r}
m2.posterior <- evidence(formula,
                         evidence.method = "CJ", # or PP
                         prior.mean = prior.mean,
                         prior.sigma = prior.sigma,
                         aux.iters = 10000, # pseudo-ERGM args
                         aux.thin = 10000,  #
                         n.aux.draws = 50,  #
                         ladder = 50,       #
                         V.proposal = 1.5,    # adjusted pseudo-posterior args
                         burn.in = 1000,      #
                         main.iters = 10000,  #
                         num.samples = 10000, # for the evidence estimation
                         estimate = "CD", # or MLE
                         seed = 1)
```
```{r}
summary(m2.posterior)

plot(m2.posterior)
```

## Model assessment

```{r}
set.seed(1)
m2.bgof <- bgof(m2.posterior,  
                sample.size = 100, # number of parameters drawn from the posterior
                aux.iters = 30000, 
                n.deg = 20,  # number of degree values to display
                n.dist = 10, # number of geodesic distance values to display
                n.esp = 7)   # number of ESP values to display

```


## Model selection

```{r}
formula <- y ~ edges +
               nodematch("Sex") +
               nodematch("Race") +
               gwesp(0.2, fixed = TRUE)

prior.mean = c(-4, 2, 2, 1)
prior.sigma = diag(2, 4)
```

### Model comparison

```{r}
m3.posterior <- evidence(formula,
                         evidence.method = "CJ", 
                         prior.mean = prior.mean,
                         prior.sigma = prior.sigma,
                         aux.iters = 10000, 
                         aux.thin = 10000,  
                         n.aux.draws = 50,  
                         ladder = 50,
                         V.proposal = 1.5,    
                         burn.in = 1000,      
                         main.iters = 10000,  
                         num.samples = 10000, 
                         estimate = "CD",
                         seed = 1)
```
```{r}
summary(m3.posterior)

plot(m3.posterior)
```


```{r}
set.seed(1)
m3.bgof <- bgof(m3.posterior,  
                sample.size = 100, 
                aux.iters = 30000,
                n.deg = 20, 
                n.dist = 10, 
                n.esp = 7)
```

```{r}
paste("M2 v M3 log evidence:", m2.posterior$log.evidence - m3.posterior$log.evidence)
```

